Title

### 2024-06-25 paper/main.tex
Updates

In this section, we develop a novel (conditional) measure of association for
mixed data by extending partial correlation to mixed data. We do this by
combining a mixed-data residualization method with a multivariate measure of
association. Given two variables $ X $ and $ Y $, and a set of conditional
variables $ \bm{Z} $, we are interested in computing an estimate of the
conditional association, $ \rho(X, Y; Z) $. Firstly, we compute the residuals
for $ x $ and $ y $ using the $ \bm{z} $ as the predictor variables. We use the
mixed data residualization approach from \citet{Ankan2023}. We consider each
variable type separately and repeat the following steps for both variables $ X
$ and $ Y $:

@@

Now, we have the two residuals and we need a measure to quantify the
association between these residuals. As the residuals are matrices, we can
treat them as sets of random variables, and in multivariate analysis canonical
correlations have been used to quantify association between two sets of
variables. Canonical correlations are generalization of correlations to
multi-dimensional variables where they try to find linear combinations of the
given sets of variables that maximize the correlation between them.

@@

	Canonical correlations try to find linear transformations $ a $ and $ b
	$ of $ \bm{U} $ and $ \bm{V} $ such that the correlations between the
	columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. The
	individual columns of $ a^T \bm{U} $ and $ b^T \bm{V} $ are
	independent. In the case when $ \rvert U \rvert = \rvert V \rvert = 1$,
	canonical correlations are equivalent to Pearson's correlation
	coefficient.

Many measures of association based on canonical correlation are also available such as:
\begin{itemize}
	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
\end{itemize}

We use Pillai's Trace in this paper. \todo[inline]{Why?}.

This measure of association has many of the desired properties:

\begin{enumerate}
	\item Is bounded between $ 0 $ and $ 1 $, with $ 0 $ representing no association and $ 1 $ representing a perfect linear relationship.
	\item Is independent of the sample size.
	\item As canonical correlation finds linear transformations, the measure is independent of the one-hot encoding scheme used for categorical variables.
	\item In the case of both continuous variables, it is equivalent to partial correlations.
	\item For both discrete variables, the measure is connected to Cramers\' V, a well know effect size measure for discrete variables.
\end{enumerate}

\section{Using measure of association for modifying DAGs}
\label{sec:modification}

>>>>

In this section, we develop a novel (conditional) measure of association for
mixed data by extending the idea of partial correlation to mixed data. We do
this by combining a mixed-data residualization method with a multivariate
measure of association. Given two variables $ X $ and $ Y $, and a set of
conditional variables $ \bm{Z} $, we are interested in computing an estimate of
the conditional association, $ \rho(X, Y; Z) $. If $ \bm{Z} = \emptyset $, we
use a constant value as $ Z $. Firstly, we compute the residuals for $ x $ and
$ y $ using the $ \bm{z} $ as the predictor variables. We use the mixed data
residualization approach from \citet{Ankan2023}. We consider each variable type
separately and repeat the following steps for both variables $ X $ and $ Y $:

@@

We now have two continuous valued residual matrices and we need a measure to
quantify the association between these residuals. In multivariate statistics,
canonical correlation based measures have been used for measuring association
between two sets of variables. We can treat the residuals as sets of variables
in case of categorical variable and use canonical corrleations to quantify the
association between them.
\todo[inline]{Justify the argument}

@@

Canonical correlations are generalization of correlations to multi-dimensional
variables. It tries to find orthogonal linear transformations $ a $ and $ b $
of $ \bm{U} $ and $ \bm{V} $ such that the correlation between the transformed
variables $ a^T \bm{U} $ and $ b^T \bm{V} $ is maximized. It gives a vector of
correlation values of size $ \min(\rvert U \rvert, \rvert V \rvert) $
representing the correlation coefficient of each column of the transformed
variables. Pearson's correlation coefficient is a special case of canonical
correlations when $ \rvert U \rvert = \rvert V \rvert = 1$.

Many measures of association based on canonical correlation have been used in
multivariate statistics with each of them having slightly different properties.
\begin{itemize}
	\item Wilks' Lambda: $ \Lambda = \prod (1 - \phi_i^2) $
	\item Roy's Largest Root: $ \theta = \phi_{\max}^2 $
	\item Pillai's Trace: $ \tau = \sum \phi_i^2 $
\end{itemize}

We use Pillai's Trace for our purpose for two main reasons: 1) It uses all the
canoncial correlation values, 2) Its interpretation is similar to correlation
coefficient, i.e., $ 0 $ signifies no association and $ 1 $ signifies perfect
linear relationship.

\todo[inline]{Add the formula for our case here}

This measure of association has many desired properties:

\begin{enumerate}
	\item Bounded between $ 0 $ and $ 1 $.
	\item Independent of sample size.
	\item Invariant to the one-hot encoding scheme used fr categorical variables.
	\item Reduces to partial correlations when both variables are continuous.
	\item In case of discrete variables, connected to Cramer\'s V, a well
		known effect size measure for discrete variables.
		\todo[inline]{Show the connection}
\end{enumerate}

\section{Using measure of association for DAG construction}
\label{sec:modification}
